Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble

### 2022-10-12 paper/paper.Rmd
update mmwy reference
add final comments from @sbfnk and @seabbs

---
title: Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble
preprint: true
author: 
  - name: Sam Abbott
    affiliation: 1
    corresponding: true
    email: sam.abbott@lshtm.ac.uk
  - name: Katharine Sherratt
    affiliation: 1
  - name: Nikos Bosse
    affiliation: 1
  - name: Hugo Gruson
    affiliation: 1
  - name:  Johannes Bracher
    affiliation: 2
  - name: Sebastian Funk
    affiliation: 1
affiliation:
  - code: 1
    address: London School of Hygiene and Tropical Medicine, London, UK
  - code: 2
    address: Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
abstract: > 
  Multi-model and multi-team ensemble forecasts have become widely used to generate reliable short-term predictions of infectious disease spread. Notably, various public health agencies have used them to leverage academic disease modelling during the COVID-19 pandemic. However, ensemble forecasts are difficult to interpret and require extensive effort from numerous participating groups as well as a coordination team. In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models. Here we used observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience to identify a set of properties present in current ensemble forecasts. We then developed a parsimonious forecast model intending to mirror these properties. We assess forecasts generated from this model in real time over six months (the 15th of January 2022 to the 19th of July 2022) and for multiple European countries. We focused on forecasts of cases one to four weeks ahead and compared them to those by the European forecast hub ensemble. We find that the surrogate model behaves qualitatively similarly to the ensemble in many instances, though with increased uncertainty and poorer performance around periods of peak incidence (as measured by the Weighted Interval Score). The performance differences, however, seem to be partially due to a subset of time points, and the proposed model appears better probabilistically calibrated than the ensemble. We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble, but more work is needed to understand the implicit epidemiological model that it represents. 
bibliography: library.bib
link-citations: true
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article
    extra_dependencies: ["float"]
    keep_tex: true
---

@@

In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysics [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
These approaches generally take the same inputs as the models they seek to emulate and then are trained based on the output from those models.
In the context of epidemiological models, non-parametric emulation has been used to allow the rapid exploration of the parameter space of complex models that would otherwise be resource-prohibitive [@Iskauskas2022-fz; @Charles2022-ds].
Replicating forecasts from resource-intensive models may also be a promising avenue of future research. 
However, as an approach, it may be less useful for resolving some of the issues of multi-team and multi-model forecasts as it does not provide interpretability, key for stakeholder take-up. Additionally, it is not clear how the methods commonly used perform out of sample, and emulating a quantile-based forecast is an unaddressed methodological challenge as far as we are aware.

In this work, we draw insights from ensemble forecasts produced and endorsed by the COVID-19 Forecast Hubs, as well as our forecasting work, to propose and evaluate a "surrogate" forecast model.
This seeks to reproduce ensemble performance by mimicking its behaviour based on a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecasts widely considered the gold standard for COVID-19 forecasting.
In addition to this, we seek to provide a simple and robust forecasting system that others can easily reuse, perhaps where a Forecast Hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and improved forecast performance.

@@

To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team. 
Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus, and asks for weekly forecasts (using MMWR epidemiological weeks i.e. Sunday to Saturday) over a time horizon of a few weeks.
Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator for other metrics such as hospitalisations [@Meakin2022-hw], and they are generally the most challenging to predict [@Sherratt2022-ue].
In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models, mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@Srivastava2020-qv; @Li2021-cj], semi-mechanistic approaches that blend both of these approaches [@Castro2021-lj; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-jn; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Karlen2020-pt; @Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].

@@

Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts.
They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions. 
In comparison to our submitted forecasts, the ensemble forecasts were less reactive to changes in trend such as from stable or reducing case incidence to increasing incidence.
On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency toward under vs over predicting. Our observations are summarised in Table \@ref(tab:observations).

```{r observations}
observations <- data.table(
  No. = 1:5,
  Observation = c(
    "Robust to daily reporting artefacts",
    "Some ability to forecast future trend changes",
    "Less reactive to apparent observed changes in trend",
    "Sharper forecasts",
    "A tendency towards under prediction"
  )
)

@@

Where $\sigma$, and $\frac{1}{\sqrt{\phi}}$ are truncated to be greater than 0 and $\beta$ is truncated to be between -1 and 1.
The Beta priors for $\xi_{+,-}$ have been chosen to be weakly informative that the reduction towards 0 growth is relatively slow.
Similarly the prior for $\beta$ has been chosen to be weakly informative that there is weak auto-correlation in differenced growth rates.
$\sigma$ has also been made weakly informative under the assumption that the potential change in growth rates in a single timestep should be relatively small.

@@

To ensure the reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with a built Docker image [@Boettiger:2015dw] (via GitHub Actions [@GitHubActionsAbout]) is provided in the code repository.
Weekly forecasts were made using `renv` and based on GitHub Actions free tier as available in 2022 to ensure they require limited compute and that our implementation is independent of local resources facilitating democratised access.
The free GitHub Actions runner we used for all forecasts was Ubuntu `20.04` based with 2 cores (x86_64), 7 GB of RAM, and 14 GB of SSD space.
The code for this analysis can be found here: https://github.com/epiforecasts/simplified-forecaster-evaluation
The code for the forecasting model defined above along with the infrastructure required to forecast using GitHub Actions can be found here: https://github.com/seabbs/ecdc-weekly-growth-forecasts
Versions archived on Zenodo are available [@REF] and [@REF].

@@

Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% prediction interval, though with a tendency to slightly over cover, with empirical coverage of `r overall_coverage[range %in% "30% interval", Surrogate]`%, `r overall_coverage[range %in% "60% interval", Surrogate]`%, `r overall_coverage[range %in% "90% interval", Surrogate]`% respectively.
The ensemble model was less well calibrated, with a tendency to under cover with empirical coverage of `r overall_coverage[range %in% "30% interval", Ensemble]`%, `r overall_coverage[range %in% "60% interval", Ensemble]`%, `r overall_coverage[range %in% "90% interval", Ensemble]`% respectively (Figure \@ref(fig:plot-coverage) a).
When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon, and then became progressively less well calibrated as the forecast horizon increased (Figure \@ref(fig:plot-coverage) a). 
In comparison, the surrogate forecast was less well calibrated than the ensemble forecast at the one-week forecast horizon with a tendency to have a larger empirical coverage than required  (Figure \@ref(fig:plot-coverage) a).
At longer horizons and narrower prediction intervals, the surrogate forecast became better calibrated though with a tendency to under cover.
This was not the case for the 90% prediction interval where the surrogate model covered more than the expected interval, for all horizons, indicating forecasts were overly uncertain for this interval regardless of the horizon.

Stratifying calibration by quantile and forecast horizon the ensemble forecast was conservative at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
This behaviour became more prominent as the forecast horizon increased.
In contrast, the surrogate forecast was generally equally well calibrated across horizons with a tendency to over cover for intervals above the median.
At longer horizons, however, quantiles below the median became liable to under cover.

@@

Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the Forecast Hub ensemble on the log scale though with greater uncertainty.
Visual differences were more marked on the natural scale with the surrogate model forecasting spuriously high peak incidence.
In a subset of example locations, we observed some variation in performance across locations, that the ensemble better-captured peak incidence, and that the surrogate model appeared biased toward overprediction.
Evaluating the relative performance of the surrogate model compared to the European Forecast Hub ensemble we found that the mean performance was substantially worse and that relative performance decreased with the forecast horizon.
The median forecast performance of the surrogate model was also worse when compared to forecasts from the ensemble though the majority of surrogate forecasts were within 50% of the performance observed for the ensemble forecast.
The difference in mean and median relative performance suggested a skewed distribution in scores, which we confirmed visually. 
This means that a relatively small fraction of forecasts were responsible for a substantial portion of the difference in performance.
Evaluating point forecast performance indicated a similar pattern of performance as that observed using the full predictive distribution though the relative performance of the surrogate model generally improved.
Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
In general, the relative performance of the surrogate model degraded as forecast horizons increased with the distribution of relative performance having an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
The Forecast Hub ensemble was poorly calibrated, particularly at longer forecast horizons and larger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
The ensemble forecast was biased towards under-prediction at short to medium forecast horizons but unbiased at longer horizons. 
In comparison, the surrogate model was biased towards over prediction and this bias increased linearly with the forecast horizon.

@@

## Conclusions {-}

@@

>>>>

---
title: Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble
params:
 preprint: TRUE
preprint: `r params$preprint`
author: 
  - name: Sam Abbott
    affiliation: 1, 2
    corresponding: true
    email: sam.abbott@lshtm.ac.uk
  - name: Katharine Sherratt
    affiliation: 1, 2
  - name: Nikos Bosse
    affiliation: 1, 2
  - name: Hugo Gruson
    affiliation: 1, 2 
  - name:  Johannes Bracher
    affiliation: 3
  - name: Sebastian Funk
    affiliation: 1, 2
affiliation:
  - code: 1
    address: The Centre for Mathematical Modelling of Infectious Diseases, London School of Hygiene & Tropical Medicine, London, UK
  - code: 2
    address: Department of Infectious Disease Epidemiology, London School of Hygiene & Tropical Medicine, London, UK
  - code: 3
    address: Chair of Statistical Methods and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
abstract: > 
  Multi-model and multi-team ensemble forecasts have become widely used to generate reliable short-term predictions of infectious disease spread. Notably, various public health agencies have used them to leverage academic disease modelling during the COVID-19 pandemic. However, ensemble forecasts are difficult to interpret and require extensive effort from numerous participating groups as well as a coordination team. In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models. Here we used observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience to identify a set of properties present in current ensemble forecasts. We then developed a parsimonious forecast model intending to mirror these properties. We assess forecasts generated from this model in real time over six months (the 15th of January 2022 to the 19th of July 2022) and for multiple European countries. We focused on forecasts of cases one to four weeks ahead and compared them to those by the European forecast hub ensemble. We find that the surrogate model behaves qualitatively similarly to the ensemble in many instances, though with increased uncertainty and poorer performance around periods of peak incidence (as measured by the Weighted Interval Score). The performance differences, however, seem to be partially due to a subset of time points, and the proposed model appears better probabilistically calibrated than the ensemble. We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble, but more work is needed to understand the implicit epidemiological model that it represents. 
bibliography: library.bib
link-citations: true
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article
    extra_dependencies: ["float"]
    keep_tex: true
   bookdown::html_document2:
    keep_toc: true        
---

@@

In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysics [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
These approaches generally take the same inputs as the models they seek to emulate and then are trained based on the output from those models.
In the context of epidemiological models, non-parametric emulation has been used to allow the rapid exploration of the parameter space of complex models that would otherwise be resource-prohibitive [@Iskauskas2022-fz; @Charles2022-ds].
These methods may be less useful for resolving some of the issues of multi-team and multi-model forecasts as they do not provide interpretability, key for stakeholder take-up. Additionally, it is not clear how these methods perform out of sample, or how they would be applied to a quantile-based forecast.

In this work, we draw insights from ensemble forecasts produced and endorsed by the COVID-19 Forecast Hubs, as well as our forecasting work, to propose and evaluate a "surrogate" forecast model.
This seeks to reproduce ensemble performance by mimicking its behaviour based on a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecasts widely considered the gold standard for COVID-19 forecasting.
Our secondary aim is to provide the basis for a robust forecasting system that others can easily reuse both in operational contexts and as a platform for future research.

@@

To understand the behaviour of the Forecast Hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team. 
Each hub targets a range of metrics, including test-positive reported cases, reported deaths, and hospitalisations; has a specific geographic focus, and asks for weekly forecasts (using MMWR epidemiological weeks i.e. Sunday to Saturday [@MMWRweek]) over a time horizon of a few weeks.
Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
Here, we focus on reported cases and primarily on the European Forecast Hub but our observations hold, in our view, across COVID-19 Forecast Hubs and to a lesser degree targets.
We focus on reported cases as these represent the most common forecast target for COVID-19 forecast models [@Nixon2022-zv], they are often of the most direct interest due to being a leading indicator for other metrics such as hospitalisations [@Meakin2022-hw], and they are generally the most challenging to predict [@Sherratt2022-ue].
In general, 5 main classes of forecast models are submitted [@Bracher2022-ug; @Cramer2022-sp], statistical forecasting models such as ARIMA models, mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@Srivastava2020-qv; @Li2021-cj], semi-mechanistic approaches that blend both of these approaches [@Castro2021-lj; @Bosse2022-hm], agent-based simulation models [@Rakowski2010-jn; @Adamik2020-if], and human insight based forecast models that may also include elements of other methods [@Karlen2020-pt; @Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].

@@

Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts.
They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions. 
In comparison to our submitted forecasts, the ensemble forecasts were less reactive to changes in trend such as from stable or reducing case incidence to increasing incidence.
On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency toward under- vs overpredicting. Our observations are summarised in Table \@ref(tab:observations).

```{r observations}
observations <- data.table(
  No. = 1:5,
  Observation = c(
    "Robust to daily reporting artefacts",
    "Some ability to forecast future trend changes",
    "Less reactive to apparent observed changes in trend",
    "Sharper forecasts",
    "A tendency towards underprediction"
  )
)

@@

Where $\sigma$, and $\frac{1}{\sqrt{\phi}}$ are truncated to be greater than 0 and $\beta$ is truncated to be between -1 and 1.
The Beta priors for $\xi_{+,-}$ have been chosen to be weakly informative that the reduction towards 0 growth is relatively slow.
Similarly the prior for $\beta$ has been chosen to be weakly informative that there is weak auto-correlation in differenced growth rates.
$\sigma$ has also been made weakly informative under the assumption that the potential change in growth rates in a single time-step should be relatively small.

@@

To ensure the reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with a built Docker image [@Boettiger:2015dw] (via GitHub Actions [@GitHubActionsAbout]) is provided in the code repository.
Weekly forecasts were made using `renv` and based on GitHub Actions free tier as available in 2022 to ensure they require limited compute and that our implementation is independent of local resources facilitating democratised access.
The free GitHub Actions runner we used for all forecasts was Ubuntu `20.04` based with 2 cores (x86_64), 7 GB of RAM, and 14 GB of SSD space.
The code for this analysis can be found here: https://github.com/epiforecasts/simplified-forecaster-evaluation
The code for the forecasting model defined above along with the infrastructure required to forecast using GitHub Actions can be found here: https://github.com/seabbs/ecdc-weekly-growth-forecasts
Versions archived on Zenodo are available [@AnalysisCode
] and [@ModelCode].

@@

Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% prediction interval, though with a tendency to be slightly underconfident, with empirical coverage of `r overall_coverage[range %in% "30% interval", Surrogate]`%, `r overall_coverage[range %in% "60% interval", Surrogate]`%, `r overall_coverage[range %in% "90% interval", Surrogate]`% respectively.
The ensemble model was less well calibrated, with a tendency to be overconfident with empirical coverage of `r overall_coverage[range %in% "30% interval", Ensemble]`%, `r overall_coverage[range %in% "60% interval", Ensemble]`%, `r overall_coverage[range %in% "90% interval", Ensemble]`% respectively (Figure \@ref(fig:plot-coverage) a).
When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon, and then became progressively less well calibrated as the forecast horizon increased (Figure \@ref(fig:plot-coverage) a). 
In comparison, the surrogate forecast was less well calibrated than the ensemble forecast at the one-week forecast horizon with a tendency to have a larger empirical coverage than required  (Figure \@ref(fig:plot-coverage) a).
At longer horizons and narrower prediction intervals, the surrogate forecast became better calibrated though with a tendency to be overconfident.
This was not the case for the 90% prediction interval where the surrogate model covered more than the expected interval, for all horizons, indicating forecasts were overly uncertain for this interval regardless of the horizon.

Stratifying calibration by quantile and forecast horizon the ensemble forecast was conservative at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
This behaviour became more prominent as the forecast horizon increased.
In contrast, the surrogate forecast was generally equally well calibrated across horizons with a tendency to be under confident for intervals above the median.
At longer horizons, however, quantiles below the median were over confident.

@@

Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the Forecast Hub ensemble on the log scale though with greater uncertainty.
Visual differences were more marked on the natural scale with the surrogate model forecasting spuriously high peak incidence.
In a subset of example locations, we observed some variation in performance across locations, that the ensemble better-captured peak incidence, and that the surrogate model appeared biased toward overprediction.
Evaluating the relative performance of the surrogate model compared to the European Forecast Hub ensemble we found that the mean performance was substantially worse and that relative performance decreased with the forecast horizon.
The median forecast performance of the surrogate model was also worse when compared to forecasts from the ensemble though the majority of surrogate forecasts were within 50% of the performance observed for the ensemble forecast.
The difference in mean and median relative performance suggested a skewed distribution in scores, which we confirmed visually. 
This means that a relatively small fraction of forecasts were responsible for a substantial portion of the difference in performance.
Evaluating point forecast performance indicated a similar pattern of performance as that observed using the full predictive distribution though the relative performance of the surrogate model generally improved.
Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
In general, the relative performance of the surrogate model degraded as forecast horizons increased with the distribution of relative performance having an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
The Forecast Hub ensemble was poorly calibrated, particularly at longer forecast horizons and larger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
The ensemble forecast was biased towards under-prediction at short to medium forecast horizons but unbiased at longer horizons. 
In comparison, the surrogate model was biased towards overprediction and this bias increased linearly with the forecast horizon.

@@

# Conclusions {-}

@@

And are archived here:

https://doi.org/10.5281/zenodo.7189308, https://doi.org/10.5281/zenodo.7189620
